---
title: "Final"
author: "Zoe Stopfer"
date: "2026-03-16"
format: html
editor: visual
---

```{r}
#| label: Reading in packages
#| message: false
#adding in packages 
library(tidyverse) 
library(here)
library(flextable)
library(janitor)
library(readxl)
library(MuMIn)
library(DHARMa)
library(ggeffects)
# reading in data 
kelp <- read_csv(here(
   "data",
    "Annual_Kelp_All_Years_20250903.csv"))
mopl <- read_csv(here(
  "data",
  "MOPL_nest-site_selection_Duchardt_et_al._2019.csv"))
Mood <- read_csv(here(
  "data",
  "JJMood4.csv"))
 
```

# Problem 1. Research Writing

## a. Transparent statistical methods

In part 1, they used a linear regressing test describing a linear relationship between wetland flooding area and precipitation. In part 2, they used a Kruskal-Wallis test since the distribution are the same in the water year classification.

## b. Figure needed

One figure that out coworker could make to accompany the Kruskal-Wallis test is a box plot. The y-axis would represent the wetland flooding area(m^2^ ).The x-axis would represent the water year classification.

## c. Suggestions for rewriting

In part 1,we retained that the precipitation doesn't predict wetland area( (linear regression, (degrees of freedom) = F-statistic R² =coefficient of determination, p = 0.11, α = significance level).

In part 2, We retained that median flooded wetland area does not differ with the water year classification ( η²= effect size) in median flooded wetland area across water year classification (wet, above normal, below normal, dry, critical drought) (Kruskal-Wallis test, '(degrees of freedom) = H-statistic p = 0.12, α = significance level). I

## d. Interpretation

One additional variable that could have influenced the flooded area of wetland was the time period. The time period is a categorical variable due to its numeric values. This variable is important because the time period could influence the amount of flooding, how much data is collected, and if it changes the data found with other tie periods.

# Problem 2. Data visualization

## a. Cleaning and Summarizing

```{r}
#| label: Clean and summarize data frame
#| message: false
kelp_clean <- kelp |> # creating a new data frame
  clean_names() |> #cleaaning names with underscores and no caps
  filter(site %in% c("IVEE", "NAPL", "CARP")) |> #filter to only have those 3 sites
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) |> # filter rows 
  mutate(site_name = case_when( # change names and put it into a new column 
  site == "NAPL" ~ "Naples",
  site == "IVEE" ~ "Isla Vista",
  site == "CARP" ~ "Carpinteria")) |> 
  mutate(site_name = as_factor(site_name), # changes "factor" values 
       site_name = fct_relevel(site_name, # appear in order of the dataset listed above
                               "Naples", "Isla Vista", "Carpinteria")) |> 
  group_by(site_name, year) |> # grouping together site and year
  summarize(mean_fronds = mean(fronds)) |> # summarize each group down to one row
  ungroup() # group by variables

```

## b. Understanding the data

The value of fronds in that "quad" in that "transect" is -99999. This value represent how there was no data collected for that particular species and the datasheet was lost as well. This information can be found in the methods tab.

## c. Visualize the data

```{r}
#| label: recreating graph
ggplot(data = kelp_clean, # read in data frame from temo_clean
       aes(x = year, # x-axis
           y = mean_fronds,#y-axis 
           color = site_name,# color and group each line by site 
           group = site_name)) + 
  geom_point() + # adding points
  geom_line() + # connect lines
  theme_minimal()+ # changing theme 
  theme(text = element_text(family = "Georgia")) +  # changing font text
  theme(legend.position = "right","top") + # move legend to right postions
  theme(plot.title = element_text(hjust = 0.5))+ # moving title position
  scale_color_manual(values = c("Naples" = "deeppink3", #add color for each site
                                "Isla Vista" = "slateblue4",
                                "Carpinteria" = "seagreen1")) +
  labs(x = "Year", # rename x, y, title, 
       y = "Mean kelp fronds per year",
       color = "Site",
       title = "Sites vary in mean kelp fronds per year") 
```

## d. Interpretation

Isla Vista has the most variable mean kelp per year. Carpinteria has the least variable mean kelp fronds per year.

# Problem 3: Data analysis

## a. Response variable

In the data set, 1s and 0s represent "yes" or "no" on whether there is a Mountain Plover nest at the site.

## b. Purpose of study

The "visual obstruction" predictor is continuous variable because we are measuring a numeric value which is the distance. As visual obstructions increases, the probability of Mountain Plover nest site use would decrease because the Mountain Plover usually breed on open flat fields, with little vegetation. This information can be found in the "Introduction", paragraph 2 of the Mountain Plover subsection.

The "distance to prairie dog colony edge" predictor is also a continuous variable since we are measuring a numeric value which is the distance. As distance to prairie dog colony edge increases, the probability of Mountain Plover site use would decrease because the Mountain Plover prefer being near prairie dog colonies since they keep the vegetation low. This information can be found in the "Introduction", paragraph 3 and 4 in the subsection regarding the black-tailed prairie dogs.

## c. Table of models

| Model Number | Visual Obstruction | Distance to Prairie Dog Colony | Model Description |
|------------------|------------------|------------------|------------------|
| 0 |  |  | no predictors (null model) |
| 1 | X | X | Saturated models |
| 2 | X |  | Visual Obstruction |
| 3 |  | X | Distance to Prairie Dog Colony |

## d. Run the models

```{r}
mopl_clean <- mopl |> # creating new data frame from mopl
  clean_names() |> #cleaaning names with underscores and no caps
  mutate(used_bin = case_match(id, 
                               "used" ~ 1,
                               "available" ~ 0)) |> # creating a new column to create 1s and 0s from the id column
  select(used_bin, edgedist, vo_nest) # selecting columns listed 
```

## e. Select the best model

```{r}
#| label: models

#model 0: null model
model0 <- glm(used_bin ~ 1,
             data = mopl_clean,
             family = "binomial")

# model 1: all predictors
model1 <- glm(used_bin ~ edgedist + vo_nest,
             data = mopl_clean,
             family = "binomial")

# model 2: Edge Distance
model2 <- glm(used_bin ~ edgedist,
             data = mopl_clean,
             family = "binomial")

# model 3: Visual Obstruction at the nest 
model3 <- glm(used_bin ~ vo_nest,
             data = mopl_clean,
             family = "binomial")

```

```{r}
#| label: AIC # testing out AIC
AICc(
  model0,
  model1,
  model2,
  model3) |> 
  arrange(AICc)
```

The best model as determined by Akaike's Information Criterion(AIC) had the visual obstruction as the predictor and the response as the used_bin ID that included yes or no(1s or 0s) on whether or not there was a nest.

## f. Check the diagnostics

```{r}
dev.off() # reset plotting
par(mar = c(4, 4, 2, 1))  # change the margins 
plot(simulateResiduals(model3)) # displaying all the code
```

## g. Visualize the model predictions

```{r}
#| label: creating model
mod_preds <- ggpredict(model3,
                       terms = "vo_nest [0:19 by = 1]") # creating new data frame from our model3 dataselt

ggplot(mopl_clean, # creating plot
       aes(x = vo_nest, # x-axis
           y = used_bin)) + # y-axis
  geom_point(size = 3, # adding dots and customizing
             alpha = 0.4,
             color = "deeppink3") +
  geom_ribbon(data = mod_preds, # creating the interval 
              aes(x = x, # x-axis, y-axis, ymin and y-max, sig level and color
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high),
              alpha = 0.4, 
              fill = "deeppink3") +
  geom_line(data = mod_preds, # adding line
            aes(x = x, # x-axis, y-axis
                y = predicted),
            linewidth = 2,# changing width
            color = "deeppink3") + # changing color
  theme_minimal() + # changing theme
  scale_y_continuous(limits = c(0, 1), # changing the y-axis/ defualt scales
                     breaks = c(0, 1)) + 
  labs(x = "Visual obstruction", # adding titles 
       y = "Probability of nest use",
       title = "Prediction of visual obstruction impact on proability of Mountain Plover nest use")
```

## h. Write a caption for your figure

**Figure 3. Maximum Likelihood Estimation(MLE) for nest use and visual obstruction.** The points at the top of the graph( in the "1"portion) represent that there is a nest. The points at the bottom of the graph (in the "0" portion) represents that there is no nest. The shaded area that follows the line represents the 95% confidence interval around the predictions and the underlying data. The curved line in the middle represents model predictions. Data from: Duchardt, Courtney; Beck, Jeffrey; Augustine, David (2020). Data from: Mountain Plover habitat selection and nest survival in relation to weather variability and spatial attributes of Black-tailed Prairie Dog disturbance \[Dataset\]. Dryad. <https://doi.org/10.5061/dryad.ttdz08kt7>

## i. Calculate model predictions

```{r}
ggpredict(model3, # make predictions 
          terms = "vo_nest [0]")
```

## j. Interpret your results

"Visual obstruction" and "distance to prairie dog colony edge" interact to predict the likelihood of a Mountain Plover nest site. As "visual obstruction" increases, the probability of a Mountain Plover nest decreases. As "distance to prairie dog colony edge" increases, the probability of a Mountain Plover nest decreases as well. To summarize, as "visual obstruction" increases so does "distance to prairie dog colony edge". The reason between obstruction and probability of Mountain Plover nest site occupancy is that the Mountain Plovers prefer places with empty landscapes and low vegetation to help avoid predictors.Prairie dogs help clear vegetation, so they would want to be closer to prairie dog colonies. At a 95% Confidence Interval( 95%CI: \[0.65, 0.80\]) the predicted probability of nest site occupancy at 0 is 0.73.

# Problem 4. Affective and exploratory visualizations

## a. Comparing visualizations

My exploratory visualizations is very simple. It has the bare bones of "attendance" and "emotion". Compared to my affective visualization, it includes other contribution factors that could have influenced my variables, such as studying an exam or being sick. Both visualizations rely on color to help distinguish the emotions. The second figure was more helpful in inspiring my affective visualization. Both visualizations follow the same trends, nothing different. During week 9 I implemented the suggestions to distinguish the days I did or didn't go to jiu jitsu but putting a "aura" around the different items. I also took the suggestion to add something when the there was a reason for not going to jiu jitsu that particular day. I thought both of these were great suggestions to make my data clearer for the reader to understand.

## b. Designing an analysis

I want to know how two variables relate to/are associated with each other so I think a Spearman correlation would be appropriate. My question would be if attendance of jiu jitsu(predictor and continuous), gives me a more positive mood( response and categorical).

## c. Check any assumptions and run your analysis

"mood_clean \<- Mood \|\> clean_names() cor.test(mood_clean$attendance,mood_clean$mood_category, \# adding in data frame and columns method = "spearman") \# this spearman test"

The test listed above would have checked my data but because of the way my data frame is set up I am unable to run a proper test since I have no numeric values in all of my data.

## d. Create a visualization

```{r}
#| message: false
mood_clean <- Mood |> 
  clean_names()
ggplot(data = mood_clean, # use my_data data frame
       aes(x = dates, # create x-axis and y-axis
           y = mood_category,
           color = mood_category)) + # color mood_category 
  geom_point() + # adding points
  labs( x = "Date", # relabelling x, y, title, and color
        y = "Mood",
        title = "Mood Throughout Time",
        color = "Mood Category") +
  theme_grey() # change theme
```

## e. Write a caption

**Figure 2. Mood Throughout Time.** Each dot represents the mood I felt at the end of the day in regards of my attendance of going to jiu jitsu. It is split up into four mood categories: red is HEP( high energy positive), green is HEP( high energy unpleasant), blue is LEP( low energy positive), and purple is LEU( low energy unpleasant).

## f. Write about your results 

My results show that when I attended jiu jitsu, I tend to have a more positive mood/ falls in a positive mood category( HEP/LEP). When I did not, my mood varied but also resulted in more negatives moods ( HEU/LEU).
