---
title: "Final"
author: "Zoe Stopfer"
date: "2026-03-16"
format: html
editor: visual
---

```{r}
#| label: Reading in packages
#| message: false
#adding in packages 
library(tidyverse) 
library(here)
library(flextable)
library(janitor)
library(readxl)
# reading in data 
kelp <- read_csv(here(
   "data",
    "Annual_Kelp_All_Years_20250903.csv"))
 
```

# Problem 1. Research Writing

## a. Transparent statistical methods

In part 1, they used a linear regressing test describing a linear relationship between wetland flooding area and precipitation. In part 2, they used a Kruskal-Wallis test since the distribution are the same in the water year classification.

## b. Figure needed

One figure that out coworker could make to accompany the Kruskal-Wallis test is a box plot. The y-axis would represent the wetland flooding area(m^2^ ).The x-axis would represent the water year classification.

## c. Suggestions for rewriting

## d. Interpretation

# Problem 2. Data visualization

## a. Cleaning and Summarizing

```{r}
#| label: Clean and summarize data frame
kelp_clean <- kelp |> # creating a new data frame
  clean_names() |> #cleaaning names with underscores and no caps
  filter(site %in% c("IVEE", "NAPL", "CARP")) |> #filter to only have those 3 sites
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) |> # filter rows 
  mutate(site_name = case_when( # change names and put it into a new column 
  site == "NAPL" ~ "Naples",
  site == "IVEE" ~ "Isla Vista",
  site == "CARP" ~ "Carpinteria")) |> 
  mutate(site_name = as_factor(site_name), # changes "factor" values 
       site_name = fct_relevel(site_name, # appear in order of the dataset listed above
                               "Naples", "Isla Vista", "Carpinteria")) |> 
  group_by(site_name, year) |> # grouping together site and year
  summarize(mean_fronds = mean(fronds)) |> # summarize each group down to one row
  ungroup() # group by variables


```

## b. Understanding the data

## c. Visualize the data

```{r}
#| label: recreating graph
#| label: ggplot
ggplot(data = kelp_clean, # read in data frame from temo_clean
       aes(x = year, # x-axis
           y = mean_fronds,#y-axis 
           color = site_name,# color and group each line by site 
           group = site_name)) + 
  geom_point() + # adding points
  geom_line() + # connect lines
  theme_minimal()+ # changing theme 
  theme(text = element_text(family = "Georgia")) +  # changing font text
  theme(legend.position = "right","top") + # move legend to right postions
  theme(plot.title = element_text(hjust = 0.5))+ # moving title position
  scale_color_manual(values = c("Naples" = "deeppink3", #add color for each site
                                "Isla Vista" = "slateblue4",
                                "Carpinteria" = "seagreen1")) +
  labs(x = "Year", # rename x, y, title, 
       y = "Mean kelp fronds per year",
       color = "Site",
       title = "Sites vary in mean kelp fronds per year") 
```
