---
title: "Final"
author: "Zoe Stopfer"
date: "2026-03-16"
format: html
editor: visual
---

```{r}
#| label: Reading in packages
#| message: false
#adding in packages 
library(tidyverse) 
library(here)
library(flextable)
library(janitor)
library(readxl)
library(MuMIn)
# reading in data 
kelp <- read_csv(here(
   "data",
    "Annual_Kelp_All_Years_20250903.csv"))
mopl <- read_csv(here(
  "data",
  "MOPL_nest-site_selection_Duchardt_et_al._2019.csv"))
 
```

# Problem 1. Research Writing

## a. Transparent statistical methods

In part 1, they used a linear regressing test describing a linear relationship between wetland flooding area and precipitation. In part 2, they used a Kruskal-Wallis test since the distribution are the same in the water year classification.

## b. Figure needed

One figure that out coworker could make to accompany the Kruskal-Wallis test is a box plot. The y-axis would represent the wetland flooding area(m^2^ ).The x-axis would represent the water year classification.

## c. Suggestions for rewriting

## d. Interpretation

# Problem 2. Data visualization

## a. Cleaning and Summarizing

```{r}
#| label: Clean and summarize data frame
#| message: false
kelp_clean <- kelp |> # creating a new data frame
  clean_names() |> #cleaaning names with underscores and no caps
  filter(site %in% c("IVEE", "NAPL", "CARP")) |> #filter to only have those 3 sites
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) |> # filter rows 
  mutate(site_name = case_when( # change names and put it into a new column 
  site == "NAPL" ~ "Naples",
  site == "IVEE" ~ "Isla Vista",
  site == "CARP" ~ "Carpinteria")) |> 
  mutate(site_name = as_factor(site_name), # changes "factor" values 
       site_name = fct_relevel(site_name, # appear in order of the dataset listed above
                               "Naples", "Isla Vista", "Carpinteria")) |> 
  group_by(site_name, year) |> # grouping together site and year
  summarize(mean_fronds = mean(fronds)) |> # summarize each group down to one row
  ungroup() # group by variables

```

## b. Understanding the data

## c. Visualize the data

```{r}
#| label: recreating graph
ggplot(data = kelp_clean, # read in data frame from temo_clean
       aes(x = year, # x-axis
           y = mean_fronds,#y-axis 
           color = site_name,# color and group each line by site 
           group = site_name)) + 
  geom_point() + # adding points
  geom_line() + # connect lines
  theme_minimal()+ # changing theme 
  theme(text = element_text(family = "Georgia")) +  # changing font text
  theme(legend.position = "right","top") + # move legend to right postions
  theme(plot.title = element_text(hjust = 0.5))+ # moving title position
  scale_color_manual(values = c("Naples" = "deeppink3", #add color for each site
                                "Isla Vista" = "slateblue4",
                                "Carpinteria" = "seagreen1")) +
  labs(x = "Year", # rename x, y, title, 
       y = "Mean kelp fronds per year",
       color = "Site",
       title = "Sites vary in mean kelp fronds per year") 
```

## d. Interpretation

# Problem 3: Data analysis

## a. Response variable

In the data set, 1s and 0s represent "yes" or "no" on whether there is a Mountain Plover nest at the site.

## b. Purpose of study

The "visual obstruction" predictor is continuous variable because we are measuring a numeric value which is the distance. As visual obstructions increases, the probability of Mountain Plover nest site use would decrease because the Mountain Plover usually breed on open flat fields, with little vegetation. This information can be found in the "Introduction", paragraph 2 of the Mountain Plover subsection.

The "distance to prairie dog colony edge" predictor is also a continuous variable since we are measuring a numeric value which is the distance. As distance to prairie dog colony edge increases, the probability of Mountain Plover site use would decrease because the Mountain Plover prefer being near prairie dog colonies since they keep the vegetation low. This information can be found in the "Introduction", paragraph 3 and 4 in the subsection regarding the black-tailed prairie dogs.

## c. Table of models

| Model Number | Visual Obstruction | Distance to Prairie Dog Colony | Model Description |
|----|----|----|----|
| 0 |  |  | no predictors (null model) |
| 1 | X | X | Satuarted model |
| 2 | X |  | Visual Obstruction |
| 3 |  | X | Distance to Prairie Dog Colony |

## d. Run the models

```{r}
mopl_clean <- mopl |> # creating new data frame from mopl
  clean_names() |> #cleaaning names with underscores and no caps
  mutate(used_bin = case_match(id, 
                               "used" ~ "1",
                               "available" ~ "0")) |> # creating a new column to create 1s and 0s from the id column
  select(used_bin,edgedist, vo_nest) # selecting columns listed 
```

## e. Select the best model

```{r}
#| label: models

#model 0: null model
model0 <- lm(used_bin ~ 1,
             data = mopl_clean)

# model 1: all predictors
model1 <- lm(used_bin ~ edgedist + vo_nest,
             data = mopl_clean)

# model 2: Edge Distance
model2 <- lm(used_bin ~ edgedist,
             data = mopl_clean)

# model 3: Visual Obstruction at the nest 
model3 <- lm(used_bin ~ vo_nest,
             data = mopl_clean)

```

```{r}
#| label: AIC
AICc(
  model0,
  model1,
  model2,
  model3) |> 
  arrange(AICc)
```

The best model as determined by Akaike's Information Criterion(AIC) had the visual obstruction as the predictor and the response as the used_bin ID that included yes or no(1s or 0s) on whether or not there was a nest.

## f. Check the diagnostics

## g. Visualize the model predictions

## h. Write a caption for your figure

## i. Calculate model predictions 

## j. Interpret your results 

# Problem 4. Affective and exploratory visualizations 

## a. Comparing visualizations 

## b. Designing an analysis

## c. Check any assumptions and run your analysis

## d. Create a visualization 

## e. Write a caption
